### 2015-09-11 notes/skinny-zone.Rmd

---
title: "Lineages around a narrow hybrid zone"
date: "`r date()`"
---

$$
\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
\DeclareMathOperator{\sgn}{sgn}
$$

```{r doc_setup, include=FALSE}
fig.dim <- 5
knitr::opts_chunk$set(fig.width=2*fig.dim,fig.align='center')
```

Suppose that heterozygotes have fitness $(1-s)$;
we know that allele frequencies around a hybrid zone behave as
$$
\begin{align} \label{eqn:pfreq}
    \partial_t p(t,x) &= \frac{\sigma^2}{2} \partial_x^2 p(t,x) - s p(t,x) (1-p(t,x)) (1-2p(t,x)) ,
\end{align}
$$
and from `lineage-movement.tex`,
if $A$ is the allele whose frequency is $p$,
an $A$-lineage follows a diffusion with generator 
$$
\begin{align*}
    G_A f(x)
    &= 
    \frac{\sigma^2}{2} (1-s(1-p(x))) \partial_x^2 f(x)
    + (2sp(x) + (1-s)) ( \partial_x \log p(x) ) \partial_x f(x) ,
\end{align*}
$$
i.e., with diffusion speed $\sigma^2 (1-s(1-p))$
and drift $(2sp(x) + (1-s)) ( \partial_x \log p(x) )$.

If we move a large distance in the direction where the $A$ allele is disadvantageous,
the $p \to 0$,
and the equation \eqref{eqn:pfreq}
becomes 
$$
\begin{align*}
    \partial_t p(t,x) &= \frac{\sigma^2}{2} \partial_x^2 p(t,x) - s p(t,x) .
\end{align*}
$$
In other words, lineages are moving by heat flow with drift.