### 2015-09-13 sims/run-sim.Rmd
added rmarkdown on using sims

---
title: "Running the simulation"
date: "`r date()`"
---

```{r doc_setup, include=FALSE}
fig.dim <- 5
knitr::opts_chunk$set(fig.height=fig.dim,fig.width=2*fig.dim,fig.align='center')
source("sim-fns.R")
```

Setting up the simulation
-------------------------

First we need to decide where the loci under selection are.
Here `qtl` is a named list, one element per chromosome,
and each element of the list is a data frame
whose columns give the name of the locus,
the selection coefficient,
and the position in Morgans.

```{r setup_qtl}
qtl=list(
         chr1=data.frame(
                    traits = c("underdominant"),
                    s = c(0.1),
                    pos = c(0.5)
                )
    )
```

Now, let's set up the population,
which means specifying the number of individuals,
the number of demes,
the width of the zone in real units,
and the dispersal SD ($\sigma$).
Our locus has a selection coefficient of `r qtl[[1]]$s`,
so the cline width will be about `r round(1/sqrt(qtl[[1]]$s),digits=1)`$\sigma$ wide;
so we'll take our zone to be four times that,
and measure distance in units of $\sigma$:
```{r setup_pop}
zone.width <- ceiling(4/sqrt(qtl[[1]]$s))  # in real units
n.ind <- 1000
n.deme <- 50
sigma <- 1 # in real units
deme.spacing <- zone.width/n.deme
deme.locs <- (seq_len(n.deme)-0.5-(n.deme/2))*deme.spacing
sigma.demes <- sigma/deme.spacing # in units of demes
```
The discretized dispersal kernel moves as $\lfloor Z+1/2 \rfloor$,
where $Z \sim N(0,\sigma_\text{demes}^2)$;
let's check this has SD close to $\sigma$:
```{r check_sigma}
zz <- floor( rnorm( 1e6, mean=1/2, sd=sigma.demes ) )*deme.spacing
sd(zz)
```
That's pretty close.

Running a simulation
--------------------

The simulation doesn't know how we're measuring things in real units,
so the sigma we pass needs to be in units of demes.
It returns all the information about the individuals,
but then we use `spBreaks()` to extract which individual is which ancestry.
```{r run_quick_sim}
sim <- sim.zone( n.ind=n.ind, n.deme=n.deme, n.gen=10, sigma=sigma.demes, quiet=TRUE )
sim.sums <- spBreaks(sim)
```

OK, now let's genotype these individuals at some loci:
```{r genotype_sim, fig.width=3*fig.dim, fig.height=1.5*fig.dim}
loci <- seq(0,1,0.05)
loci.cols <- rainbow(12)[cut(abs(loci-0.5),breaks=min((length(loci)-1)/2,8))]
info <- data.frame(
               chr=1,
               bp=floor(loci*1e6),
               pos=loci
           )
genotypes <- geno.pop( sim.sums$ind.ancest, list(chr1=loci) )
```

Now, let's
compute ancestry frequencies,
and plot the resulting clines:
```{r genotype_freqs, fig.width=3*fig.dim, fig.height=1.5*fig.dim}
freqs <- do.call( cbind, tapply( 1:ncol(genotypes), sim$deme, function (kk) {
                    rowMeans(genotypes[,kk])
                } ) )
matplot(deme.locs, t(freqs), type='l', col=loci.cols, xlab="spatial position", ylab="ancestry frequency")
legend("topleft",lty=1,legend=loci,col=loci.cols,title="locus position")
```

We can also look at the local frequencies of the diploid genotypes at each locus:
```{r diploid_freqs}
diploid.freqs <- lapply( seq_along(loci), function (locus) {
                do.call( rbind, tapply( 1:ncol(genotypes), sim$deme, function (kk) {
                    as.vector( tabulate( 1+genotypes[locus,kk], nbins=3 ) )
                } ) )
           } )
# one end
matplot(diploid.freqs[[1]],type='l',main=paste("locus at",loci[1]),lty=1,col=c('red','black','green'))
legend('topright',lty=1,col=c('red','black','green'),legend=c("homozygote","heterozygote","homozygote"))
# the selected locus
sel.locus <- match( qtl$chr1$pos, loci )
matplot(diploid.freqs[[sel.locus]],type='l',main=paste("selected locus at",loci[sel.locus]),lty=1,col=c('red','black','green'))
legend('topright',lty=1,col=c('red','black','green'),legend=c("homozygote","heterozygote","homozygote"))
# the other end
matplot(diploid.freqs[[length(loci)]],type='l',main=paste("locus at",loci[length(loci)]),lty=1,col=c('red','black','green'))
legend('topright',lty=1,col=c('red','black','green'),legend=c("homozygote","heterozygote","homozygote"))
```