### 2015-10-15 notes/brownian-local-times.Rmd
.

---
title: "Brownian local times"
date: "`r date()`"
---

$$
\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
\DeclareMathOperator{\sgn}{sgn}
$$

```{r doc_setup, include=FALSE}
fig.dim <- 5
knitr::opts_chunk$set(fig.width=2*fig.dim,fig.align='center')
```

Killing at Brownian local times
===============================

It states in [Takacs](https://projecteuclid.org/euclid.aoap/1177004703) that 
if $\ell(x)$ is the local time at $x$ for Brownian motion run for time 1 and begun at 0,
i.e.,
$$
\begin{aligned}
    \ell(x) = \lim_{\epsilon \to 0} \frac{1}{\epsilon}
        \text{measure}\{t \,:\, x \le B_t \le x+\epsilon, \, 0 \le t \le 1\}
\end{aligned}
$$
then
$$
\begin{aligned}
    \P\{ \ell(x) \le u \} = 2 \Phi(|x|+u)-1,
\end{aligned}
$$
where $\Phi(x)=\P\{B_1 \le x\}$ is the Gaussian CDF.
Differentiating this, $\P\{\ell \in du\} = \partial_u \P\{\ell \le u\}$, so
$$
\begin{aligned}
    \P\{ \ell(x) = 0 \} &= 2 \Phi(|x|)-1 \\
    \P\{ \ell(x) \in du \} &= 2 \phi(|x|+u) ,
\end{aligned}
$$
where $\phi$ is the Gaussian density.

Let's check this:
```{r sim_rw, cache=TRUE}
nwalks <- 1e4
ntimes <- 1e3
# want total time to be 1, so:
eps <- 1/sqrt(ntimes)
rws <- matrix( rnorm(nwalks*ntimes)*eps, nrow=ntimes, ncol=nwalks )
rws <- rbind(0,apply(rws,2,cumsum))
matplot(seq(0,1,length.out=ntimes),rws[,1:20],type='l', ylab='position')
```
Here we've simulated `r nwalks` random walks for a total of `r ntimes` steps,
and scaled steps to have this be a total Brownian time of 1.
Let's check;
this should just be a standard Gaussian:
```{r final_dist}
xh <- hist(rws[ntimes,],breaks=50,main="final position")
lines( xh$mids, sum(xh$counts)*diff(pnorm(xh$breaks)) )
```

Now, let's check on the local time formula.
The number of steps the random walk spends at position $k$ on a grid with spacing $\epsilon$
and a time step of \gamma
should be $\epsilon \times \ell(k\epsilon) / \gamma$.
First, the local time at 0:
```{r local_times_0, cache=TRUE, depends="sim_rw"}
xh <- hist(colSums(abs(rws)<=1.5*eps)/ntimes/(3*eps),main="local time at 0", breaks=30)
lines( xh$mids, sum(xh$counts)*diff( 2*pnorm(xh$breaks)-1 ), col='red' )
```
What about the local time at other locations?
Note there is an atom at 0 now
(that is depicted by the dots in the figures below),
and that the discretization could trip us up;
the colors (red/green/blue) use the (mid/upper/lower) points of the interval that we count in.
```{r local_times, cache=TRUE, depends="sim_rw"}
layout(matrix(1:4,nrow=2))
for (loc in c(0.1, 0.2, 0.4, 0.6)) {
    lts <- colSums(abs(rws-loc)<=1.5*eps)/ntimes/(3*eps)
    zeros <- (lts==0)
    pred.zero <- 2*pnorm(abs(loc)+c(0,1.5*eps,-1.5*eps))-1
    xh <- hist(lts[!zeros],main=paste("local time at",loc), breaks=30, ylim=range(c(0,pred.zero*nwalks)))
    lines( xh$mids, sum(!zeros)*diff( 2*pnorm(abs(loc)+xh$breaks)-1 ), col='red' )
    points(rep(0,4), c(sum(zeros),pred.zero*nwalks), col=1:4, pch=20, cex=2)
    # lower bound
    lines( xh$mids, sum(!zeros)*diff( 2*pnorm(abs(loc+1.5*eps)+xh$breaks)-1 ), col=3 )
    # upper bound
    lines( xh$mids, sum(!zeros)*diff( 2*pnorm(abs(loc-1.5*eps)+xh$breaks)-1 ), col=4 )
}
```
Clearly, the discretization is still tripping us up a bit,
but it's pretty good.

Its Laplace transform
=====================

Now suppose that something happens at rate $r$ relative to the local time at $x$
(e.g., recombination to the other background).
Then the chance that this has not happened by time 1 is
$$
\begin{aligned}
\zeta(x,r) &:=
\E\left[ e^{-r\ell(x)} \right] \\
&=
2\Phi(|x|)-1 + 
\int_0^\infty e^{-rt} 2 \phi(|x|+t) dt \\
&=
2\Phi(|x|)-1 + 
e^{r|x|} \int_{|x|}^\infty e^{-rt} 2 \phi(t) dt \\
&=
2\Phi(|x|)-1 + 
2 e^{r|x|} \int_{|x|}^\infty \frac{1}{\sqrt{2\pi}} \exp\left( - rt - t^2/2 \right) dt \\
&=
2\Phi(|x|)-1 + 
2 e^{r|x|+r^2/2} \int_{|x|}^\infty \frac{1}{\sqrt{2\pi}} \exp\left( - (t+r)^2/2 \right) dt \\
&=
2\Phi(|x|)-1 + 
2 e^{r|x|+r^2/2} \left( 1-\Phi(|x|+r) \right) .
\end{aligned}
$$
At $r=0$ this equals 1, as it should, since it's the chance that nothing happens.
For large $x$, since $1-\Phi(x) \sim e^{-x^2/2}/\sqrt{2\pi}x$,
this is approximately 
$1 + 2 \exp( r|x| + r^2/2 - (|x|+r)^2/2 )$,
which goes to 1, as it should.

Let's check that, too: why not?
```{r killing, cache=TRUE, depends="sim_rw"}
loc <- 0.1
zeta <- function (x,r) {
    2*pnorm(abs(x)) - 1 + 2 * exp( r*abs(x) + r^2/2 + pnorm(abs(x)+r,lower.tail=FALSE,log.p=TRUE) )
}
compare.rates <- as.data.frame( t( sapply( c(0.25,0.5,1,1.5,2,2.5,3), function (rate) {
        visits <- colSums(abs(rws-loc)<=1.5*eps)
        killed <- ( rbinom(ncol(rws),size=visits,prob=rate/ntimes/(3*eps)) == 0 )
        c( 
          rate=rate,
          simulation=mean(killed), 
          theory=zeta(loc,rate),
          SE=sqrt(mean(killed)*(1-mean(killed))/nwalks)
          )
    } ) ) )
with(compare.rates, {
         plot( rate, theory, type='l', xlab="recomb rate", ylab="prob of recomb", col='red', ylim=c(0,1) ) 
         points( rate, simulation, pch=20 )
         arrows( x0=rate, y0=simulation-2*SE, y1=simulation+2*SE, angle=90, code=3 )
    } )
```

This is very steep at $x=0$.
Notice that as $x$ goes to 0,
$$
\begin{aligned}
    \lim_{x \to 0} \zeta(x,r) 
    &= 2 e^{r^2/2} (1-\Phi(r)) \\
    &= \frac{2}{r\sqrt{2\pi}} + O(r^{-3}) .
\end{aligned}
$$
so that the difference in the left and right limits is
$$
\begin{aligned}
    \lim_{x \to 0-} p(x,r) 
    -\lim_{x \to 0+} p(x,r) 
    &= 
    \frac{1}{2}+\frac{2}{r\sqrt{2\pi}} 
    - \frac{1}{2}-\frac{2}{r\sqrt{2\pi}} 
      +O(r^{-3}) \\
    &=
    \frac{4}{r\sqrt{2\pi}}
      +O(r^{-3}) 
\end{aligned}
$$

Let's also look at the slope of $p$ at $x=0$.
Since $\partial_x |x| = \sgn(x)$,
$$
\begin{aligned}
    \partial_x \zeta(x,r)
    &=
    2 \sgn(x) \left(
    \phi(|x|)
    + r e^{r|x|+r^2/2} (1-\Phi(|x|+r))
    - e^{r|x|+r^2/2} \phi(|x|+r)
      \right)
\end{aligned}
$$
and so
$$
\begin{aligned}
    -\partial_x p(x)
    &=
    \phi(|x|)
    + r e^{r|x|+r^2/2} (1-\Phi(|x|+r))
    - e^{r|x|+r^2/2} \phi(|x|+r) \\
    &\to
    \frac{1}{\sqrt{2\pi}}(1-e^{r^2/2}e^{-r^2/2})
    + r e^{r^2/2} (1-\Phi(r))  \qquad \text{as } x \to 0 \\
    &=
    r e^{r^2/2} (1-\Phi(r)) \\
    &=
    \frac{1}{\sqrt{2\pi}} + O(\frac{1}{r}) .
\end{aligned}
$$

Also note that for large $|x|$, by the asymptotics for $1-\Phi(x)$,
$$
\begin{aligned}
    \zeta(x,r) = 1 - \sqrt{\frac{2}{\pi}} e^{-x^2/2} \left( \frac{r}{x+r} \right) .
\end{aligned}
$$

What this says about clines
===========================

The above restricted to standard Brownian motion on one unit of time.
Really, we will want to use this across $\tau$ generations,
modeling a random walk with variance $\sigma^2$ per generation.

The key quantity we want is the probability that a Brownian motion with variance $\sigma^2$ killed at rate $r$
with respect to its local time at $x$ was still alive at time $\tau$.
Denote this by
$$
\begin{aligned}
    \zeta(x,r,\tau,\sigma).
\end{aligned}
$$
By Brownian scaling,
since $B_{ct}$ has the same distribution as $\sqrt{c}B_t$,
$$
\begin{aligned}
    \zeta(x,r,\tau,\sigma) 
    &=
    \zeta(x,r\tau,1,\sqrt{\tau}\sigma) \\
    &=
    \zeta(x/\sqrt{\tau\sigma^2},r\tau) ,
\end{aligned}
$$
where
$$
\begin{aligned}
    \zeta(x,r)
    &=
    1 - 2(1-\Phi(|x|)) + 
    2 e^{r|x|+r^2/2} \left( 1-\Phi(|x|+r) \right) .
\end{aligned}
$$

**Cline flattening:**
immediately, this tells us how a cline gets wider as time progresses:
for a fixed $r$, the width grows as $\sqrt{\tau}$.
It also tells us that clines with the same $r\tau$ have the same shape,
after rescaling the width by $\sqrt{\tau}$.
Also, this tells us we only have a cline for $r$ of order at least $1/\tau$:
for smaller $r$, the locus is still completely linked to the selected locus,
and hence a step function.

More generally, here's what $\zeta$ looks like:
```{r plot_zeta, fig.height=2*fig.dim}
layout(1:2)
xvals <- seq(0,5,length.out=150)
rvals <- seq(0,10,length.out=100)
zetavals <- outer( xvals, rvals, zeta )
matplot(xvals,zetavals,type='l',xlab="x",ylab=expression(zeta),col=rainbow(length(rvals)))
lrvals <- seq(1,length(rvals),length.out=6)
legend("bottomright",lty=1,col=rainbow(length(rvals))[lrvals],legend=paste("r=",round(rvals[lrvals],digits=2)))
matplot(rvals,t(zetavals),type='l',xlab="r",ylab=expression(zeta),col=rainbow(length(xvals)))
lxvals <- seq(1,length(xvals),length.out=6)
legend("bottomright",lty=1,col=rainbow(length(xvals))[lxvals],legend=paste("x=",round(xvals[lxvals],digits=2)),bg='white')
```

Now, let $p(x,r,\tau,A)$ denote the probability that a lineage begun at $x<0$
at a location at recombination distance $r$ from the selected site
inherits from the left-hand side of the barrier
(i.e., an ancestor on $(-\infty,0)$)
$\tau$ generations ago.
(Note that $r$ should be scaled by the density and fecundity of heterozygotes in the cline --
see below.)
For an individual to the left of the barrier ($x<0$),
the probability of inheriting from the left of the barrier at $\tau$
is the probability that no recombination at the barrier has occurred,
plus one-half the probability that recombination in a heterozygote has occurred,
i.e., $\zeta +(1-\zeta)/2=(1+\zeta)/2$.
If $x>0$ this is the one-half times probability a recombination in a heterozygote has occurred,
i.e., $(1-\zeta)/2$.
Combining these,
the probability of ending up on the left-hand side is
$$
\begin{aligned}
    p(x,r,\tau,A) = 
    \frac{1}{2} \left\{ 1 - \sgn(x) \zeta(x,r,\tau) \right\} .
\end{aligned}
$$

Here's the cline shape:
```{r show_cline, cache=TRUE}
pcline <- function (x,r,tau=1,sigma=1) {
    (1/2)*( 1 - sign(x)*zeta(x=x/sqrt(tau*sigma^2),r=r*tau) )
}
xx <- seq(-15,15,length.out=500)
for (tau in c(5,50)) {
    plot(xx, pcline(xx,r=1), type='n', 
            xlab="distance from cline center", ylab="prob inherited from the left", 
            main=as.expression(substitute(tau==mytau,list(mytau=tau))) )
    rr <- seq(0,0.2,length.out=11)
    pp <- sapply( rr, function (r) { pcline(xx,r=r,tau=tau) } )
    matplot(xx,pp,type='l',lty=1,col=rainbow(length(rr)),add=TRUE)
    legend("topright",lty=1,col=rainbow(length(rr)),legend=paste("r=",rr))
}
```

Adjusting for cline width
=========================

Above, we were approximating the time that a reflecting lineage
spends in a small interval near zero
by the local time at zero multiplied by the width of the interval.
In fact, when the lineage is at $x$, 
the lineage recombines to the other background
in a given generation
with probability equal to the recombination rate 
multiplied by the probability of being found in a heterozygote.
If $p_{AB}(x)$ is the local fraction of heterozygotes,
and heterozygotes have relative fecundity $1-s$,
then the probability of being in a heterozygote is
$(1-s)p_{AB}(x)/((1-s)p_{AB}(x)+2p_{AA}(x))$;
if diploid genotypes are assorted randomly,
then $p_{AB} = 2p(1-p)$ and $p_{AA}=p^2$, so the probability that an $A$ lineage
passes through a heterozygote in a given generation is
$$
\begin{align*}
f_H(x) 
&= 
\frac{ (1-s)p(x)(1-p(x)) }{ p(x)^2 + (1-s) p(x) (1-p(x)) }
&= 
\frac{ (1-s)(1-p(x)) }{ p(x) + (1-s) (1-p(x)) } \\
&= 
\frac{ (1-s)(1-p(x)) }{ 1 - s (1-p(x)) } .
\end{align*}
$$

Bazykin's approximation to a cline is
$$
p(x) = \frac{1}{2} (1+\tanh(-2x\sqrt{s}/\sigma))
$$
Here's what that looks like:
```{r tanh_cline, cache=TRUE}
tanh_cline <- function (x,s,sigma=1) {
    (1/2)*(1+tanh(-2*x*sqrt(s)/sigma))
}
fH <- function (x,s,pfn=tanh_cline,...) {
    px <- pfn(x,s=s,...)
    ( (1-s)*(1-px) )/( 1 - s*(1-px) )
}
plot(xx, tanh_cline(xx,s=0.04), type='l' )
```

A lineage does something a bit more complicated,
but let's just run a random walk with drift equal to the logarithmic derivative of $p(x)$,
i.e.,
$$
\begin{align*}
    \partial_x \log(p(x)) 
    &= -\frac{2\sqrt{s}}{\sigma} 
        \frac{1-\tanh^2(-2x\sqrt{s}/σ)}{1+\tanh(-2x\sqrt{s}/σ)} .
\end{align*}
$$

```{r sim_lineage, cache=TRUE, depends=c("tanh_cline","show_cline","killing")}
drift <- function (x,s,sigma=1) {
    -(2*sqrt(s)/sigma) * (1-tanh(-2*x*sqrt(s)/sigma)^2) / (1+tanh(-2*x*sqrt(s)/sigma))
}
nwalks <- 1e4
ntimes <- 3e3
total.time <- 10
s <- 0.1
dt <- total.time/ntimes
# put the noise in rws
rws <- rbind( rnorm(nwalks), # initial conditions
             matrix( rnorm(nwalks*(ntimes-1)), nrow=(ntimes-1), ncol=nwalks ) )
# and then integrate over it
for (k in 2:ntimes) {
    rws[k,] <- rws[k-1,]+sqrt(dt)*rws[k,]+dt*drift(rws[k-1,],s)
}
layout( t(1:2), widths=c(3,1) )
matplot(seq(0,total.time,length.out=ntimes),rws[,1:80],type='l', ylab='position', xlab='time', ylim=range(rws) )
legend("topright",lwd=2,legend=c("prob of het","cline","drift"),col=c("black","green","red"))
plot( (-1)*drift(xx,s=s), xx, col='red', lwd=2, type='l', ylim=range(rws), ylab='', xlab='drift or prob' )
lines( tanh_cline(xx,s=s), xx, col='green', lwd=2 )
lines( (1-s)*(1-tanh_cline(xx,s=s))/(1-s*(1-tanh_cline(xx,s=s))), xx, lwd=2 )
```
The cline is in green and the local probability of being in a heterozygote is in black;
the drift (multiplied by $-1$) is in red.

Let's just check that's doing the right thing:
```{r sim_lineage_check, cache=TRUE, depends="sim_lineage"}
xbreaks <- seq(-3,3,length.out=100)
xmids <- xbreaks[-1]-diff(xbreaks)/2
dX <- apply(rws,2,diff)
Xmeans <- tapply( dX, cut(rws[-nrow(rws),],breaks=xbreaks), mean )
Xvar <- tapply( dX, cut(rws[-nrow(rws),],breaks=xbreaks), var )
layout(1:2)
plot(xmids,Xmeans/dt,ylab="mean dX")
lines(xmids, drift(xmids,s), col='red', lwd=2)
plot(xmids,Xvar/dt,ylab="variance dX")
abline(h=1,col='red',lwd=2)
```

**Comparison to zero-width cline:**
Let's start a bunch of random walks sorta far from the cline center,
and look at the distribution of the time spend in heterozygotes.
```{r sim_het_time, cache=TRUE, depends="sim_lineage"}
nwalks <- 1e4
ntimes <- 3e3
total.time <- 100
s <- 0.1
dt <- total.time/ntimes
# put the noise in rws
rws <- rbind( -10, # initial conditions
             matrix( rnorm(nwalks*(ntimes-1)), nrow=(ntimes-1), ncol=nwalks ) )
# and then integrate over it
for (k in 2:ntimes) {
    rws[k,] <- rws[k-1,]+sqrt(dt)*rws[k,]+dt*drift(rws[k-1,],s)
}
layout( t(1:2), widths=c(3,1) )
matplot(seq(0,total.time,length.out=ntimes),rws[,1:80],type='l', ylab='position', xlab='time', ylim=range(rws) )
legend("topright",lwd=2,legend=c("prob of het","cline","drift"),col=c("black","green","red"))
plot( (-1)*drift(xx,s=s), xx, col='red', lwd=2, type='l', ylim=range(rws), ylab='', xlab='drift or prob' )
lines( tanh_cline(xx,s=s), xx, col='green', lwd=2 )
lines( (1-s)*(1-tanh_cline(xx,s=s))/(1-s*(1-tanh_cline(xx,s=s))), xx, lwd=2 )
```
Here is the distribution of $T=\int_0^{`r total.time`} f_H(X_t) dt$:
```{r dist_het_time, cache=TRUE, depends="sim_het_time"}
het.time <- dt*colSums( fH(rws,s=s) )
layout(1:2)
hist( het.time, breaks=50, xlab="time in hets", freq=FALSE, main="histogram of T" ) 
hist( log(het.time), breaks=50, xlab="log(time in hets)", freq=FALSE , main="histogram of log(T)") 
```

Brownian motion with drift above zero
=====================================

A better model for the process might be Brownian motion
that, when above zero, has negative drift,
and is killed at some rate.
In other words, 
if $dX_t = -\theta(X_t) dt + dB_t$,
where $\theta(x)=\theta$ for $x \ge 0$, then
we want to know
$$\begin{aligned}
    u(x;\lambda,\theta):=\E^x \left[ \exp\left( - \lambda \int_0^\infty X_t dt \right) \right] .
\end{aligned}$$
(*Small problem:*  this isn't guarenteed to be on the correct size of zero when we want it to.)
Well, by Girsanov's theorem, we know that
$\E[f(X_t)] = \E[f(B_t)\zeta(t)]$
where
$$\begin{aligned}
\zeta(t) = \exp\left[ \int_0^t \theta(B_u) dB_t - \frac{1}{2} \int_0^t \theta(B_u)^2 du \right] 
\end{aligned}$$
and since by Tanaka's formula,
$$\begin{aligned}
\theta \max(0,B_t) - \max(0,B_0) = \int_0^t \theta(B_u) du + \theta \ell_t(0) ,
\end{aligned}$$
thus
$$\begin{aligned}
\zeta(t) = \exp\left[ \theta \max(0,B_t) - \theta \max(0,B_0) - \theta \ell_t(0) - \frac{1}{2} \int_0^t \theta(B_u)^2 du \right] 
\end{aligned}$$

More than we need to know
-------------------------

According to Karatzas & Shreve (1984)
(with their $\delta=\theta_1=0$),
if $dX_t = -\theta(X_t) dt + dB_t$,
where $\theta(x)=\theta$ for $x \ge 0$, then
if we define
$$
\begin{aligned}
h(s;x,\mu) = \frac{|x|}{\sqrt{2\pi s^3}} \exp\left[-\frac{(x+\mu s)^2}{2s}\right] \qquad x>0 ,
\end{aligned}
$$
(this is the density of the first time a BM with drift $\mu$ started at $x$ hits 0),
then
$$
\begin{aligned}
p_t(x,z) 
&= \P\{ X_t = z \mid X_0 = x \}\\
&= \begin{cases}
    2 \int_0^\infty \int_0^t h(t-\tau;b-z,0) h(\tau;x+b,\theta) d\tau db
        & \qquad x \ge 0, \quad z\le 0 \\
    2 \int_0^\infty \int_0^t e^{2\theta z} h(t-\tau;b,0) h(\tau;x+b,\theta) d\tau db & \\
    \qquad {} + \frac{1}{\sqrt{2\pi t}} 
        \left\{ 
            \exp\left[ -\frac{ (x-z+\theta t)^2 }{2t} \right]
            +\exp\left[ -\frac{ (x+z-\theta t)^2 }{2t} - 2 \theta x \right]
        \right\}
        & \qquad x \ge 0, \quad z > 0 .
\end{cases}
\end{aligned}
$$
Also, $p_t(x,z;\theta)=p_t(-x,-z;-\theta)$ gets the same expressions for $x<0$.