### 2015-10-22 sims/run-sim-Alisa.Rmd
got .rmd to work-ish!

---
title: "Running simulations by Alisa"
author: Alisa
date: "`r date()`"
output: html_document
---

```{r doc_setup, include=FALSE}
fig.dim <- 5
knitr::opts_chunk$set(fig.height=fig.dim,fig.width=2*fig.dim,fig.align='center')
source("sim-fns.R")
```

Setting up the simulation
-------------------------
To simulate, we read in files from sim-fns.R
```{r load_fns}
	source("sim-fns.R")
```

Start off with simulation of 50 generation zone with 50 demes of 500 individuals each and sigma of 1:
```{r set_deme_parameters}
	zone_age	= 50
	ndemes 		= 50
	deme_size 	= 500
	ninds 		= ndemes*deme_size
	SIGMA 		= 1
```
note that zone_age above can be a vector of times.

Now to define selection. Here we selection against heterozygotes to be strong (0.5).
```{r set_selection}
	S  = 0.5
	qtl=list(chr1=data.frame(traits=c("underdominant"), s = S,pos=c(0.5)))
```
`qtl` is a named list, one element per chromosome,
and each element of the list is a data frame
whose columns give the name of the locus,
the selection coefficient,
and the position in Morgans.

Running the simulation
-------------------------

This is a two-step process. First we simulate genotypes:
```{r run_sim_step1}
	sims.s0.1 = lapply(zone_age,sim.zone,n.ind=ninds,n.deme=ndemes,sigma=SIGMA)
	#This can take a while, 
```
The output is: a list with hierarchy: [[zone_age]][[c("inds","sp.inds","deme","pars","QTL")]]
	where:
		"inds" is a list of length ninds, such that [[ind]][[chr]][[X1/X2]][[data.frame where col1 = coordinates of rec., col2 = ancestor ID]]
		"sp.inds" is a character vector of length ninds, giving the ancestry given an ancestor ID
		"deme" is an integer vector of length ninds, giving the deme number of an ancestor, given their ancestor ID
		"pars" is an integer that gives the number of generatsion of the simulation
		"QTL" returns the QTL vector used in the simulation

The second step is to parse the simulated genotypes:
```{r run_sim_step2}
	sims.sums = lapply(sims.s0.1,spBreaks)
```
The output here is a list: [[zone_age]][[c("ind.ancestry","ancest.prop")]]
		"ind.ancestry" is a list of length ninds: [[ind][[chr]]][[X1/X2]][[data.frame where col1 = (chunk)starts, col2 = (chunk)stops, col3 = sp2(T/F)]]
		"ancest.prop" is a ninds*2 matrix giving the ancestry proportion of each individual on each of their two copies of the chromosome.

This resulting files can get quite large, but we've saved them here
```{r save_file,cache=TRUE}
	save(sims.s0.1,file="sims_sums_s0.5_sigma1_50by500_tau50.Robj")
```

Getting relevant statistics
-------------------------
Let's start by defining a vector of loci at which we will measure things:

```{r focal_loci}
	loci = list(seq(0.5,1,0.05))
```

Now, we get frequencies at each of the sites in the above vector, 
first by getting genotypes at all loci, then using those to calculate an allele frequency for each deme:
```{r get_frequencies}
	genotypes = lapply(sims.sums,function(Z){data.frame(do.call(cbind,lapply(Z$ind.ancest, geno.ind, loci)))})
	freqs = lapply(genotypes,function(X){apply(X,1,function(Z){tapply(Z,cut(1:ncol(X),breaks=seq(0,ncol(X),deme_size)),mean)/2})})
	xx <- (1:ndemes)-0.5-ndemes/2
		matplot(xx,freqs[[1]],col=rainbow(20),lty=1,type="l", main="50 generations",ylab="freq",xlab="deme")
		legend('bottomright',col=rainbow(20),lty=1,legend=loci[[1]],cex=0.75)
		matpoints(-xx,pp,col=rainbow(20),lty=3,type="l")

```
(`genotypes` is a list of nloci*ninds dataframes where each entry is the number of ancestry B alleles in an individual at the given locus.)
Each line in the plot represents a locus a given distance from the selected locus.

We can compare these frequencies to predictions made by theory :

```{r peters_predictions}
	zeta <- function(x,r){
	2*pnorm(abs(x)) - 1 + 2 * exp(r*abs(x) + r^2/2 + pnorm(abs(x)+r,lower.tail=FALSE,log.p=TRUE))	
	}

	pcline <- function(x,r,tau=zone_age,sigma=SIGMA){
	(1/2)*(1-sign(x)*zeta(x=x/sqrt(tau*sigma^2),r=r*tau))	
	}

	rr=(loci[[1]]-0.5)*SIGMA/sqrt(S) # loci at which frequency is measured
	xx <- (1:ndemes)-0.5-ndemes/2	 # deme positions
	pp <- do.call(cbind,lapply(rr,function(R){pcline(xx,r=R,tau=zone_age,sigma=SIGMA)})) #allele frequency

		matplot(xx,freqs[[1]],col=rainbow(20),lty=1,type="l", main="50 generations",ylab="freq",xlab="deme")
		legend('bottomright',col=rainbow(20),lty=1,legend=loci[[1]],cex=0.75)
		matpoints(-xx,pp,col=rainbow(20),lty=3,type="l")

```

Next, we think about tract lengths in these populations. The following generates the distribution of lengths 
of continuous ancestry B to the right of the specified locus (`focal_sites`) (in this case the selected locus). 
Note that if the specified locus is of ancestry A, then the length of ancestry B is 0. The first function finds
and measures the chunk in an individual that contains the focal site. The second function generates a 
distribution for each deme.

```{r tract_length_functions}
	focal_sites = lapply(qtl,function(Z){Z$pos})

	get.interval.size = function(IND_DATA=sims.sums[[1]]$ind.ancest[[1]],CHR=1,POS=0.5,ancA = TRUE){
		#return the interval containing focal site for a individual	
		chunk = sapply(IND_DATA[[CHR]],function(X){diff(as.numeric(X[which(X$starts<POS & X$stops>POS),1:2]))})	
		identity = sapply(IND_DATA[[CHR]],function(X){X[which(X$starts<POS & X$stops>POS),3]})
		replace(chunk,which(identity==ancA),0)
	}

	get.deme.chunks = function(IND_DATA=sims.sums[[1]]$ind.ancest, DEME = 1, CHR=1,POS=0.5,ancA = TRUE){
		#get distribution of chunks within a deme
		INDS = IND_DATA[which(sims.s0.1[[1]]$deme==DEME)]
		intervals = do.call(rbind,lapply(INDS,get.interval.size,CHR=CHR,POS=POS,ancA=ancA)) 	
		return(intervals)
	}

```
the output of get.deme.chunks is a nind*2 matrix, where each column is a chromosome. To run:

```{r get_tract_lengths}
	tracts = lapply(1:ndemes,function(X){get.deme.chunks(DEME=X,ancA=F)})

		#function to increase transparency of colors
		add.alpha <- function(col, alpha=1){
		if(missing(col))
		stop("Please provide a vector of colours.")
		apply(sapply(col, col2rgb)/255, 2, 
		function(x) 
		rgb(x[1], x[2], x[3], alpha=alpha)) 
		}

	transparent_rainbow = add.alpha(rainbow(ndemes),0.75)

		D = 1; hist(tracts[[D]], border="black",breaks=seq(0,1,0.05),ylim=c(0,100),xlab="length",main="Distribution of tracts")

		for(D in seq(1,ndemes,5)){
		hist(tracts[[D]], col=transparent_rainbow[D],border=NA,breaks=seq(0,1,0.05),ylim=c(0,50),add=T)	
		}
```
To get LD in these populations, we define `get.genotype` to get the genotypes at two loci in individuals, and `get.LD` to calculate LD in a deme:
```{r LD_functions}
 	get.genotype = function(IND_DATA=sims.sums[[1]]$ind.ancest[[1]],CHR=1,POS1=0.5,POS2=0.5){
	site1 = sapply(IND_DATA[[CHR]],function(X){X[which(X$starts<POS1 & X$stops>POS1),3]})	
	site2 = sapply(IND_DATA[[CHR]],function(X){X[which(X$starts<POS2 & X$stops>POS2),3]})
	return(rbind(site1,site2))
	}

	get.LD = function(IND_DATA= sims.sums[[1]]$ind.ancest,DEME=1,CHR=1,POS1 = 0.5, POS2=0.1){
	INDS = IND_DATA[which(sims.s0.1[[1]]$deme==DEME)]
	genotypes = do.call(cbind,lapply(INDS,get.genotype,CHR=CHR,POS1=POS1,POS2=POS2))
	freqB1 = length(which(genotypes[1,]))/length(genotypes[1,])
	freqB2 = length(which(genotypes[2,]))/length(genotypes[2,])
	freqB12 = length(which(genotypes[1,] & genotypes[2,]))/length(genotypes[1,])
	LD = freqB12 - freqB2*freqB1
	return(LD)
}

```
We can just calculate LD over demes and loci pairs. Here's an example of LD between pairs of loci where one member of the
pair is selected on. This gives a plot of decay in population LD as we get further away from the center of the zone.

```
LD_matrix = do.call(cbind, lapply(seq(0.51,0.99,0.05),function(Z){sapply(1:ndemes,get.LD,POS1=0.5,POS2=Z,CHR=1,IND_DATA=sims.sums[[1]]$ind.ancest)}))
	matplot(xx,LD_matrix,type="l",col=rainbow(20),xlim=c(-10,10),lty=1)

```

Simulate a realistic population (with Populus as a model)
-------------------------
We propose that the density of Populus along a one-dimensional transect is one every 10 meters. Additionally, based on a
literature search that suggests most pollen only makes it to the nearest neighbor and seeds do not disperse much, with 
potentially a long tail, we propose a sigma of 100m. Furthermore, based on HZ width of 20km for most of the geome, we 
propose that width at selected loci may be 1km, giving a selection coefficient of 0.01 (based on width = 2*sigma/sqrt(S)). 
We space demes ever 1km for 100km, with 100 individuals per deme (given above density of 1/10 inds/meter). In units of demes, sigma is 10.

```{r set_populus_parameters,cache=TRUE}
	zone_age	= 10
	ndemes 		= 100
	deme_size 	= 100
	ninds 		= ndemes*deme_size
	SIGMA 		= 0.1
	S 			= 0.01
	qtl=list(chr1=data.frame(traits=c("underdominant"), s = S,pos=c(0.5)))

	sims.populus_like.10gens = lapply(zone_age,sim.zone,n.ind=ninds,n.deme=ndemes,sigma=SIGMA)
	sims.sums.populus_like.10gens = lapply(sims.populus_like.10gens,spBreaks)
	save(sims.populus_like.10gens,file="simspopulus_like.10gens.Robj")
```