### 2015-10-30 notes/rough-calculations.Rmd
.

---
title: "Clines from PDE"
date: "`r date()`"
---

$$
\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\R}{\mathbb{R}}
\DeclareMathOperator{\sgn}{sgn}
\newcommand{\grad}{\nabla}
$$

```{r doc_setup, include=FALSE}
fig.dim <- 5
knitr::opts_chunk$set(fig.width=2*fig.dim,fig.height=fig.dim,fig.align='center')
library(ReacTran)
source("tran.1D.R") # fix for log.A
```

*Useful functions used below:*
```{r setup_fns}
source("cline-fns.R")
```

In the following, $x$ is the distance from the cline center,
$r$ is the recombination rate to the selected locus,
$\tau$ is the time since secondary contact.
The two alleles are $A$ and $B$.
We would like to know:

1.  What is the width of the cline at a linked, neutral locus?
    How long does it take to flatten out?
    How far from the cline do you have to go to find mostly one type?

2.  How long are the haplotypes around the rare $B$ alleles on the $A$ side?
    This should be constant in time.

3.  How long, and how common, are chunks of $B$ haplotype *not* linked to $B$ alleles on the $A$ side?
    These will accumulate as time progresses.

4.  How does neighborhood size affect the results?
    Does the mean cline, averaged across low-population-density simulations,
    match the cline obtained by a high population density, i.e., the PDE?

5.  What about the variance??