### 2015-12-03 notes/time-scales.Rmd
fixed yinit at zero

---
title: "Time scales for cline flattening"
date: "`r date()`"
---

```{r doc_setup, include=FALSE}
fig.dim <- 5
knitr::opts_chunk$set(fig.width=2*fig.dim,fig.height=fig.dim,fig.align='center')
library(ReacTran)
library(Matrix)
source("tran.1D.R") # fix for log.A
source("cline-fns.R")
```

Here we would like to explore what happens over various time scales.
We'll do:

-  tens of generations
-  hundreds of generations
-  thousands of generations

At each time scale, we'll look at:

-  linked clines
-  ancestry disequilibrium
-  haplotype lengths

Tens of generations
===================

First, the setup:
```{r setup_tens, cache=TRUE}
s <- 0.1
tt <- seq(0,40,length.out=21)
xgrid <- setup.grid.1D(x.up=-20.5, x.down=20.5, N=41)
rr <- seq(0,0.2,length.out=21)
rcols <- rainbow(length(rr)*1.2)[seq_along(rr)]
fwds.soln <- forwards_pde(s=s,times=tt,grid=xgrid)
pfun <- cline_fun(fwds.soln)
```
We'll solve it up until time `r max(tt)` and on a grid of `r xgrid$N` sites from `r min(xgrid$x.mid)` to `r max(xgrid$x.mid)`.

Here's the behaviour of the selected site that we use to describe the behaviour of the other sites:
```{r fwds_soln_plot, cache=TRUE, depends="setup_tens"}
image(fwds.soln, legend=TRUE, zlim=c(0,1), yaxt='n', main="selected site", ylab="space")
add_grid_axis(xgrid)
```

Linked clines
-------------

```{r tens_linked_clines, cache=TRUE, depends="setup_tens"}
linked.solns <- lapply( rr, function (r) {
            forwards_backwards_pde(s=s,times=tt,grid=xgrid,r=r,log.p=TRUE,fwds.soln=fwds.soln)
        } )
linked.clines <- lapply( linked.solns, cline_from_soln, grid=xgrid )
```

Here are linked clines for a few time steps:
```{r plot_tens_linked_clines, cache=TRUE, depends="tens_linked_clines"}
for (plot.t in c(4,10,20,40)) {
    k <- which.min(abs(tt-plot.t))
    lclines <- sapply( linked.clines, cline_interp, t=tt[k] )
    matplot( xgrid$x.mid, lclines, type='l', ylab="proportion", xlab='space', 
            main=sprintf("t=%0.0f",tt[k]),
           lty=1, col=rcols )
    legend("topright", lty=1, col=rcols, legend=sprintf("r=%0.2f",rr) )
}
```

Ancestry disequilibrium
-----------------------

Now we'll look at ancestry disequilibrium across the region.

```{r tens_ld_solutions, cache=TRUE}
rr <- seq(-0.2,0.2,length.out=21)
rcols <- rainbow(length(unique(abs(rr)))*1.2)[match(abs(rr),unique(abs(rr)))]
rpairs <- as.matrix(subset(expand.grid(r1=rr,r2=rr),r1<r2))
LD.solns <- parallel::mclapply( 1:nrow(rpairs), function (k) 
               forwards_backwards_haplotypes(s=s, times=tt, xgrid=xgrid, rgrid=list(x.mid=rpairs[k,],N=2), fwds.soln=fwds.soln), 
           mc.cores=parallel::detectCores() )
LD.values <- lapply( LD.solns, AD_from_soln, xgrid=xgrid )
```

Here are heatmaps of ancestry LD at a few different times and locations
The selected site is in the center.
```{r tens_ld_heatmaps, cache=TRUE, depends="tens_ld_solutions", fig.width=3*fig.dim}
plot.x <- c(0,5,10)
these.x <- 
layout(t(1:3))
for (plot.t in c(4,10,20,40)) {
    image( 
}
```