### 2015-12-05 notes/computing-AD.Rmd
begun computing AD

---
title: "Measures of correlation in ancestry"
date: "`r date()`"
---

```{r doc_setup, include=FALSE}
fig.dim <- 5
knitr::opts_chunk$set(fig.width=2*fig.dim,fig.height=fig.dim,fig.align='center')
```

Goal
=====

Compute something like correlation in ancestry between nearby markers,
in different locations,
as a measure of typical haplotype length in the (spatial and/or genomic) region.

**Nomenclature:** Yaniv followed John in calling it "ancestry disequilibrium".
This admirably follows population genetics tradition
in choosing a name that is in comprehensible to outsider.
I propose calling it something that uses "correlation" instead of "disequilibrium"?
(*Todo:* figure out what the measures below are correlations of, and say that.)

Suggestion
==========

Fix spatial locations $x$, $y$, and $z$.
and loci $i$ and $j$.
Suppose we sample individuals at $x$, $y$, and $z$,
and then in each of these samples, at each locus, pick a random allele;
let $(X_i,X_j)$ be the alleles sampled thus at $x$, and likewise for $Y$ and $Z$.
(If we have phased data, instead pick a random haplotype.)
Then, define
\begin{align} \label{eqn:Dx_defn}
    D_x(i,j) = \P\{ X_i=Y_i\neq Z_i \and X_j=Y_j\neq Z_j\} - \P\{ X_i=Y_i \neq Z_i \} \P\{ X_j=Y_j \neq Z_j \} .
\end{align}

Now, how to estimate this?

Suppose $I$ has alleles $a$ and $A$, and $J$ has alleles $b$ and $B$;
and suppose that at location $x$ we have sampled $n_x$ individuals,
of which $n_x(aa,bb)$ have diploid genotype $a/a$, $b/b$, etcetera.
Define $p_x = n_x/n$, and so forth.
The first term is
\begin{align} \label{eqn:Dx_defn}
    \P\{ X_i=Y_i\neq Z_i \and X_j=Y_j\neq Z_j\} 
\end{align}