### 2015-12-11 notes/computing-AD.Rmd
estimating AD

---
title: "Measures of correlation in ancestry"
date: "`r date()`"
---

```{r doc_setup, include=FALSE}
fig.dim <- 5
knitr::opts_chunk$set(fig.width=2*fig.dim,fig.height=fig.dim,fig.align='center')
```

Goal
=====

Compute something like correlation in ancestry between nearby markers,
in different locations,
as a measure of typical haplotype length in the (spatial and/or genomic) region.

**Nomenclature:** Yaniv followed John in calling it "ancestry disequilibrium".
This admirably follows population genetics tradition
in choosing a name that is incomprehensible to outsider.
I propose calling it something that uses "correlation" instead of "disequilibrium"?
(*Todo:* figure out what the measures below are correlations of, and say that.)

Suggestion
==========

Fix spatial locations $x$, $y$, and $z$.
and loci $i$ and $j$.
Suppose we sample individuals at $x$, $y$, and $z$,
and then in each of these samples, at each locus, pick a random allele;
let $(X_i,X_j)$ be the alleles sampled thus at $x$, and likewise for $Y$ and $Z$.
(If we have phased data, instead pick a random haplotype.)
Then, define
\begin{align} \label{eqn:Dx_defn}
    D_x(i,j) = \P\{ X_i=Y_i\neq Z_i \and X_j=Y_j\neq Z_j\} - \P\{ X_i=Y_i \neq Z_i \} \P\{ X_j=Y_j \neq Z_j \} .
\end{align}

Now, how to estimate this?

Suppose $I$ has alleles $a$ and $A$, and $J$ has alleles $b$ and $B$;
and suppose that at location $x$ we have sampled $n_x$ individuals,
of which $n_x(aa,bb)$ have diploid genotype $a/a$, $b/b$, etcetera.
First, compute the two-locus haplotype probabilities:
\begin{align}
    p_x(a,b)
    &=
    \frac{1}{n_x} \left( 
        n_x(aa,bb) 
        + \frac{1}{2} n_x(aa,bB)
        + \frac{1}{2} n_x(aA,bb)
        + \frac{1}{4} n_x(aA,bB)
     \right)
\end{align}
and so forth.
We can get these by defining $p_x(a,b) = n_x(a,b)/n_x$, where expected haplotype counts are defined by:
\begin{align}
    \begin{bmatrix} n(a,b) \\ n(A,b) \\ n(a,B) \\ n(A,B) \end{bmatrix}
    &=
    \begin{bmatrix}     1    & \frac{1}{2} &    0     & \frac{1}{2} & \frac{1}{4} &    0        &    0      &    0        &     0   \\ % ab
                        0    & \frac{1}{2} &    1     &    0        & \frac{1}{4} & \frac{1}{2} &    0      &    0        &     0   \\ % Ab
                        0    &     0       &    0     & \frac{1}{2} & \frac{1}{4} &    0        &    1      & \frac{1}{2} &     0   \\ % aB
                        0    &     0       &    0     &    0        & \frac{1}{4} & \frac{1}{2} &    0      & \frac{1}{2} &     1      % AB
    \end{bmatrix}
    \begin{bmatrix} n(aa,bb) \\ n(aA,bb)  \\ n(AA,bb) \\ n(aa,bB)  \\ n(aA,bB)   \\   n(AA,bB) \\ n(aa,BB) \\ n(aA,BB)   \\ n(AA,BB) \end{bmatrix}
\end{align}
Denote by $Q$ the matrix above, and note that columns of $Q$ sum to 1.

The first term is estimated by
\begin{align} 
    \P\{ X_i=Y_i\neq Z_i \and X_j=Y_j\neq Z_j\} 
    &=
    p_x(a,b) p_y(a,b) p_z(A,B)
    + p_x(A,b) p_y(A,b) p_z(a,B)
      \\ & \qquad {}
    + p_x(a,B) p_y(a,B) p_z(A,b)
    + p_x(A,B) p_y(A,B) p_z(a,b)
\end{align}
(In this ordering, reversing the vector complements it, so in R this would be `p[,x]*p[,y]*rev(p[,z])`.)

We would like to estimate the probability in the second term
by independent, i.e., *distinct* samples,
so it is a bit more involved.
First, suppose that we draw two individuals from our sample from $x$
*without replacement*, and pick a haplotype in each.
If the first genotype is $(aa,bB)$, for instance, then the probability of obtaining a $b$ on the second draw is
$$ \frac{n_x(aa,bb) + n_x(aA,bb) + \frac{1}{2} (n_x(aa,bB)-1) + \frac{1}{2} n_x(aA,bB) }{ n_x } = \frac{ n_x(\cdot,b) - 1/2 }{ n_x-1 } . $$
Similar calculations show that 
the probability that we get an $a$ at the first locus in the first draw and a $b$ at the second locus in the second draw is
$$
    \tilde p_x(a,b) = \frac{n_x}{n_x-1} p_x(a,\cdot) p_x(\cdot,b) - \frac{1}{n_x-1} \sum_g p_x(g) H_g(b)^2,
$$
where the sum is over diploid genotypes $g$, and
$H_g(b)$ is the first row of $H_g$, obtained by multiplying $Q$ on the left by a $2 \times 4$ matrix:
\begin{align}
H_g = \begin{bmatrix}
    1 & 1 & 1 & 1/2 & 1/2 & 1/2 & 0 & 0 & 0 \\
    0 & 0 & 0 & 1/2 & 1/2 & 1/2 & 1 & 1 & 1 
\end{bmatrix} .
\end{align}
Using this definition, the second term is then estimated as
\begin{align} 
    \P\{ X_i=Y_i \neq Z_i \} \P\{ X_j=Y_j \neq Z_j \}
    &=
    \tilde p_x(a,b) \tilde p_y(a,b) \tilde p_z(A,B)
    + \tilde p_x(A,b) \tilde p_y(A,b) \tilde p_z(a,B)
      \\ & \qquad {}
    + \tilde p_x(a,B) \tilde p_y(a,B) \tilde p_z(A,b)
    + \tilde p_x(A,B) \tilde p_y(A,B) \tilde p_z(a,b)
\end{align}